---
layout: post
title: "Make XKCD Style Graphs in R"
author: [yihui]
categories: [Fun, Font]
tags: [XKCD, Humor Sans, Cairo]
reviewer: []
---
{% include JB/setup %}

Fonts play an important role in graphics. This article was motivated from a [StackOverflow
post](http://stackoverflow.com/q/12675147/559676), and shows how to draw [xkcd](http://xkcd.com)
style graphs in R using the [**Cairo** package](http://cran.r-project.org/package=Cairo). Before
get started, you need to install the Humor Sans font (please Google it yourself).

## Set up fonts for R

After the font has been installed, we can tell **Cairo** to use it:

```{r set-font, tidy=FALSE}
library(Cairo)
CairoFonts(
  regular = "Humor Sans:style=Regular",
  bold = "Humor Sans:style=Bold",
  italic = "Humor Sans:style=Italic",
  bolditalic = "Humor Sans:style=Bold Italic,BoldItalic"
)
```

## Draw it

Now we use the `CairoPNG()` device to save a plot. Because I'm using **knitr**, I only need to
specify the chunk option `dev='CairoPNG'` instead of really opening the device in the code below.

```{r draw-xkcd, dev='CairoPNG', dependson='set-font', fig.height=5}
# you may need CairoPNG() and dev.off() around this chunk
set.seed(20121108)
x = seq(1, 10, .1); r = function() runif(length(x), -.05, .05)
y1 = sin(x) + r(); y2 = cos(x) + r()
# draw the lines and texts
par(mar = c(2, .1, .1, .1))
plot.new()
plot.window(range(x), c(-1.5, 1))
grid()
matplot(x, cbind(y1, y2, y2), type = 'l', lty = 1,
        col = c('black', 'white', 'red'), lwd = c(3, 15, 3), add = TRUE)
axis(1, c(2, 5, 6, 9), c("YARD", "STEPS", "DOOR", "INSIDE"), lwd = 0, line=-1)
text(7, -1.25, "A SIN AND COS CURVE", cex=1.5)
lines(x, -1.5 + runif(length(x), -.005, .005), xpd = NA, lwd=2)
```

There are other ways to use additional fonts in R graphics, such as the **tikzDevice** and
**extrafonts** package; the former requires knowledge of LaTeX, and the latter is easier to use.

## Acknowledgments

I thank Mark Bulling for the answer under the SO post. I used base R graphics instead of
**ggplot2** as Mark did because I think the pen-and-paper model of base R is more natural for
drawing such a plot (the grammar of graphics seems to be overkill).

## Session info

Since I'm not entirely sure about how **Cairo** performs under Windows; here is my R session info:

```{r sessionInfo}
print(sessionInfo(), locale = FALSE)
```
