---
layout: post
title: "Demonstration of the Central Limit Theorem"
author: [yulijia]
categories: [Animation, Probability]
tags: [Central Limit Theorem, Poisson distribution, Cauchy distribution]
reviewer: [yihui]
animation: true
---
{% include JB/setup %}

In [Probability Theory](http://en.wikipedia.org/wiki/Probability_theory), the [Central Limit
Theorem (CLT)](http://en.wikipedia.org/wiki/Central_limit_theorem) states that, given certain
conditions, the mean of a sufficiently large number of independent random variables, each with a
well-defined mean and well-defined variance, will be approximately normally distributed.

As is shown in [The Bean Machine and the Central Limit
Theorem](http://vis.supstat.com/2013/04/bean-machine/), the central limit theorem has a number of
variants. This article shows you as long as the conditions of the Central Limit Theorem (CLT) are
satisfied, the distribution of the sample mean will be approximate to the Normal distribution when
the sample size n is large enough, no matter what is the original distribution.

In the [**animation** package](http://yihui.name/animation), there is a function named `clt.ani()`.
It shows the distribution of the sample mean when the sample size $$n$$ grows up. The normality
test `shapiro.test()` is performed to check whether the sample mean follows the normal
distribution.

## Classical Central Limit Theorem

With the parameter `FUN` in the function `clt.ani()` you can select distribution which will be
shown in the animation. Here is the example with [Poisson distribution](http://en.wikipedia.org/wiki/Poisson_distribution).

```{r poisson-clt, fig.show='animate', interval=.5, fig.height=8, fig.width=6}
library(animation)
ani.options(interval = 0.5)
par(mar = c(3, 3, 1, 0.5), mgp = c(1.5, 0.5, 0), tcl = -0.3)
lambda=4
f = function(n) rpois(n, lambda)
clt.ani(FUN = f, mean = lambda, sd = lambda)
```

## The Central Limit Theorem does not work

The [Cauchy distribution](http://en.wikipedia.org/wiki/Cauchy_distribution) is an example of a
distribution which has no mean, variance or higher moments defined. The Central Limit Theorem describes the relation of a sample mean to the population mean. So we can't apply Central Limit Theorem to the distribution.

```{r cauchy-clt, fig.show='animate', interval=.5, fig.height=8, fig.width=6}
ani.options(interval = 0.5)
par(mar = c(3, 3, 1, 0.5), mgp = c(1.5, 0.5, 0), tcl = -0.3)
f = function(n) rcauchy(n,location=0,scale=2)
clt.ani(FUN = f)
```
