---
layout: post
title: "Cheat Sheets for Plotting Symbols and Color Palettes"
author: [rcore, yihui, yulijia]
categories: [Base Graphics, Color]
tags: [plotting symbols, color palettes, graphical parameters, RColorBrewer]
reviewer: []
---
{% include JB/setup %}

This cheat sheet shows a few commonly used plotting symbols and color palettes in R. Now you do not
need to memorize the facts like "19 is the big solid dot and 21 is an open circle that can have a
different background color" -- just bookmark this cheat sheet. In fact it was motivated by Dr
[Rafael Irizarry](http://rafalab.jhsph.edu/), who printed the color palettes on a piece of paper
and pinned that to the wall in his office.

## Plotting symbols (pch)

There are many plotting symbols in the **graphics** package. We can set the `pch` parameter to set
the symbols. See `?points` for more information.

Below is a figure containing the plot symbols from `pch = 0` to `25` along with some
character-based plot symbols. We can use, for example, `plot(x, y, pch = 3)` for plus signs (`+`)
in a scatterplot.

```{r pch, fig.height=5, echo=FALSE}
# taken from example(points)
pchShow = function(extras = c("*",".", "o","O","0","+","-","|","%","#"), cex = 3,
                   col = "red3", bg = "gold", coltext = "brown", cextext = 1.2) {
  nex <- length(extras)
  np  <- 26 + nex
  ipch <- 0:(np-1)
  k <- floor(sqrt(np))
  dd <- c(-1,1)/2
  rx <- dd + range(ix <- ipch %/% k)
  ry <- dd + range(iy <- 3 + (k-1)- ipch %% k)
  pch <- as.list(ipch) # list with integers & strings
  if(nex > 0) pch[26+ 1:nex] <- as.list(extras)
  par(mar=rep(0,4))
  plot(rx, ry, type = "n", axes  =  FALSE, xlab = "", ylab = "")
  abline(v = ix, h = iy, col = "lightgray", lty = "dotted")
  for(i in 1:np) {
    pc <- pch[[i]]
    ## 'col' symbols with a 'bg'-colored interior (where available) :
    points(ix[i], iy[i], pch = pc, col = col, bg = bg, cex = cex)
    if(cextext > 0)
      text(ix[i] - 0.3, iy[i], pc, col = coltext, cex = cextext)
    }
  }
pchShow()
```

## Color palettes

The default color palette in R:

```{r default-pal, fig.width=5, fig.height=5}
(pal = palette())
par(mar=rep(0,4))
pie(rep(1, length(pal)), labels=sprintf('%d (%s)', seq_along(pal),pal), col=pal)
```

Below are the color palettes in **RColorBrewer** (if you do not want to use **ggplot2**, you should
consider using this package to generate colors for your plots; just do not use the meaningless
`rainbow()` palette):

```{r brewer-pal, fig.height=9}
library(RColorBrewer)
par(mar=c(0,4,0,0))
display.brewer.all()
# generate 8 colors from the Set2 palette
brewer.pal(8, 'Set2')
```

In R, the function `colors()` returns a vector of `r length(colors())` color names. When you really
need to specify the color by its name, you are recommended to check out this nice [color
chart](http://research.stowers-institute.org/efg/R/Color/Chart/ColorChart.pdf) created by Earl F.
Glynn.
