---
layout: post
title: "Least Squares Fitting"
author: [yulijia]
categories: [Animation, Linear Models]
tags: [Least Squares]
reviewer: [yihui]
animation: true
---
{% include JB/setup %}

[Least squares](http://en.wikipedia.org/wiki/Least_squares) fitting is a mathematical procedure for
finding the best-fitting curve to a given set of points by minimizing the sum of the squares of the
offsets ("the residuals") of the points from the curve. *Least squares* means that the overall
solution minimizes the sum of the squares of the errors made in the results of every single
equation.

The function `least.squares()` in the [**animation** package](http://yihui.name/animation) has
provided an illustration of least squares in univariate linear regression.

let's assume that the relationship between x andy is a linear one.

Let $$y=ax+b$$ be the eauation of the best fit line to the data. We need to determine the valuses
of both the slop $$a$$ and intercept $$b$$. Assuming each data point carries eauql weight, each
$$y_{i}$$ point has exactly the same actual error asscociated with it, we find $$a$$ and $$b$$ by
minimizing the sum of the squares of the deviations of the actual values of $$y_{i}$$ from the
line's calcualted value of y.

The formulas for $$a$$ and $$b$$ are:

$$a=SLOPE=\frac{n\sum{x_{i}y_{i}}-\sum {x_{i}}\sum{y_{i}}}{n\sum{x_{i}^{2}}-(\sum{x_{i}})^2}$$

$$b=INTERCEPT=\frac{\sum{x_{i}^{2}}\sum{y_{i}}-\sum{x_{i}}\sum{y_{i}}}{n\sum{x_{i}^{2}}-(\sum{x_{i}})^2}$$

Here is the simple demonstration of the meaning of least squares in univariate linear regression.

## With slope changing


```{r slope-changing, fig.show='animate', interval=.3}
library(animation)
ani.options(interval = 0.3)
par(mar = c(4, 4, 0.5, 0.1), mgp = c(2, 0.5, 0), tcl = -0.3)
## slope changing
least.squares()
```

## With intercept changing
```{r intercept-changing , fig.show='animate', interval=.3}
library(animation)
ani.options(interval = 0.3)
par(mar = c(4, 4, 0.5, 0.1), mgp = c(2, 0.5, 0), tcl = -0.3)
# slope changing
least.squares(ani.type = "intercept")
```
