---
layout: post
title: "A Cheat Sheet for Plotting Symbols and Color Palettes in R"
author: [yihui, yulijia]
categories: [Graphical Parameters]
tags: [Plotting Symbols, Color Palettes]
reviewer: []
---
{% include JB/setup %}

This cheat sheet shows how many plottins symbols and color palettes in R. Now you do not need ask
others the plotting symbols and color palettes questions again, just save this cheat sheet to your
favorite folder, everything gets ready:)

## Plotting symbols (pch)

There’s a rich variety of plotting symbols in the **graphics** package. We can set `pch` parameter
to change plotting symbols. See `?points` for more information.

This script creates a figure containing the standard plot symbols along with some character-based
plot symbols.

```{r pch, fig.height=5}
# taken from example(points)
pchShow = function(extras = c("*",".", "o","O","0","+","-","|","%","#"), cex = 3,
                   col = "red3", bg = "gold", coltext = "brown", cextext = 1.2) {
  nex <- length(extras)
  np  <- 26 + nex
  ipch <- 0:(np-1)
  k <- floor(sqrt(np))
  dd <- c(-1,1)/2
  rx <- dd + range(ix <- ipch %/% k)
  ry <- dd + range(iy <- 3 + (k-1)- ipch %% k)
  pch <- as.list(ipch) # list with integers & strings
  if(nex > 0) pch[26+ 1:nex] <- as.list(extras)
  par(mar=rep(0,4))
  plot(rx, ry, type = "n", axes  =  FALSE, xlab = "", ylab = "")
  abline(v = ix, h = iy, col = "lightgray", lty = "dotted")
  for(i in 1:np) {
    pc <- pch[[i]]
    ## 'col' symbols with a 'bg'-colored interior (where available) :
    points(ix[i], iy[i], pch = pc, col = col, bg = bg, cex = cex)
    if(cextext > 0)
      text(ix[i] - 0.3, iy[i], pc, col = coltext, cex = cextext)
    }
  }
pchShow()
```

## Color palettes

The default color palette in R:

```{r default-pal, fig.width=5, fig.height=5}
(pal = palette())
par(mar=rep(0,4))
pie(rep(1, length(pal)), labels=sprintf('%d (%s)', seq_along(pal),pal), col=pal)
```

The color palettes in **RColorBrewer**:

```{r brewer-pal}
library(RColorBrewer)
par(mar=c(0,4,0,0))
display.brewer.all()
```

We can set R colors by name, or by index in the `colors()` vector. Function `colors()` returns a
vector of 657 color names in R.

Still need more? See [color
chart](http://research.stowers-institute.org/efg/R/Color/Chart/ColorChart.pdf).
