---
layout: post
title: "Mathematical Annotation in R"
author: [rcore, yulijia, kbroman, kushey]
categories: [Base Graphics]
tags: [mathematical annotation, Greek letters]
reviewer: [yihui]
---
{% include JB/setup %}

Want to write mathematical symbols and expressions in R graphics? You can use an R `expression()`
instead of normal text, e.g. `plot(1:10, main = expression(alpha + beta))`. Below is a demo that
shows you everything about plotting math in R (it was written by the R Core Team; see `?plotmath`
for details):

```{r demo-plotmath, eval=FALSE}
demo(plotmath)
```
```{r read-plotmath, cache=FALSE, include=FALSE}
read_demo('plotmath', package='graphics', labels='plotmath')
```
```{r plotmath, echo=FALSE, results='hide', dpi=90, crop=TRUE}
```

## Combining expressions and text

If you want to combine multiple mathematical expressions with text, use `paste()` _inside_
`expression()`, as in the following.

```{r math-text, fig.width=5, fig.height=4, tidy=FALSE}
par(mar = c(4, 4, 2, 0.1))
plot(rnorm(100), rnorm(100),
  xlab = expression(hat(mu)[0]), ylab = expression(alpha^beta),
  main = expression(paste("Plot of ", alpha^beta, " versus ", hat(mu)[0])))
```

Finally, if we want to include variables from an R session in mathematical expressions, and
substitute in their actual values, we can use `substitute()`.

```{r math-text-sub, fig.width=6, fig.height=4, tidy=FALSE}
par(mar = c(4, 4, 2, 0.1))
x_mean <- 1.5
x_sd <- 1.2
hist(rnorm(100, x_mean, x_sd),
  main = substitute(
    paste(X[i], " ~ N(", mu, "=", m, ", ", sigma^2, "=", s2, ")"),
    list(m = x_mean, s2 = x_sd^2)
  )
)
```
