---
layout: post
title: "Mathematical Annotation in R"
author: [rcore, yulijia, kbroman]
categories: [Base Graphics]
tags: [mathematical annotation, Greek letters]
reviewer: []
---
{% include JB/setup %}

Want to write mathematical symbols and expressions in R graphics? You can use an R `expression()`
instead of normal text, e.g. `plot(1:10, main = expression(alpha + beta))`. Below is a demo that
shows you everything about plotting math in R (it was written by the R Core Team; see `?plotmath`
for details):

```{r demo-plotmath, eval=FALSE}
demo(plotmath)
```
```{r read-plotmath, cache=FALSE, include=FALSE}
read_demo('plotmath', package='graphics', labels='plotmath')
```
```{r plotmath, echo=FALSE, results='hide', dpi=90, crop=TRUE}
```

## Combining expressions and text

If you want to combine multiple mathematical expressions with text, use `paste` _inside_
`expression`, as in the following.

```{r math-text, fig.width=5, fig.height=4, tidy=FALSE}
par(mar=c(4, 4, 2, .1))
plot(rnorm(100), rnorm(100),
  xlab=expression(hat(mu)[0]), ylab=expression(alpha^beta),
  main=expression(paste("Plot of ", alpha^beta, " versus ", hat(mu)[0])))
```
