---
layout: post
title: "Simulation of Coin Flipping with R"
author: [yulijia]
categories: [Animation, Probability]
tags: [Coin flipping, Bernoulli trial]
reviewer: [yihui]
animation: true
---
{% include JB/setup %}

In the [**animation** package](http://yihui.name/animation), the function `flip.coin()` provides a simulation to the process of flipping coins and computes the frequencies for 'heads' and 'tails'. Coin flipping is a famous [Bernoulli trial](http://en.wikipedia.org/wiki/Bernoulli_trial). When you flip a coin, there are two possible outcomes: heads and tails. A fair coin has the probability of 'heads' 0.5 by definition.

## Head or tail

In this case, we toss a fair coin 100 times.

```{r head-or-tail, fig.show='animate', fig.height=5, interval=.3}
library(animation) 
ani.options(nmax = 100, interval=.3)
par(mar = c(2, 4, 2, 2))
flip.coin(bg = "yellow")
```

## Coin stands

In this case, we set the coin can stands when flipping.

```{r coin-stands, fig.show='animate', fig.height=5, interval=.3}
ani.options(nmax = 100, interval=.3)
par(mar = c(2, 4, 2, 2))
flip.coin(faces = c("Head", "Stand", "Tail"), type = "n", prob = c(0.45, 0.1, 0.45), col = c(1, 2, 4))
```
